# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r, echo=TRUE}
data<-read.csv("activity.csv",stringsAsFactors=FALSE)
```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
library(reshape2)
datesteps <- melt(data[,1:2], id=c("date"),na.rm=TRUE)
mean_datesteps<-dcast(datesteps, date~variable,mean)
hist(mean_datesteps$steps)
mean(mean_datesteps$steps)   # the mean total number of steps taken per day
median(mean_datesteps$steps) # the median total number of steps taken per day
```

## What is the average daily activity pattern?
```{r, echo=TRUE}
library(reshape2)
intervalsteps <- melt(data[,c(1,3)], id=c("interval"),na.rm=TRUE)
mean_intervalsteps<-dcast(intervalsteps,interval~variable,mean)
plot(x=mean_intervalsteps$interval,y=mean_intervalsteps$steps,type="l")
# the maximum number of steps
mean_intervalsteps$interval[
  which(mean_intervalsteps$steps==max(mean_intervalsteps$steps))] 
```


## Imputing missing values
```{r, echo=TRUE}
sum(is.na(data$steps)==TRUE) # the total number of missing values
# Create a New data set
filldata <- data
# Devise a strategy "mean for that 5-minute interval"
# for filling in all of the missing values in the dataset.
for (i in 1:length(filldata$steps)) {
    if (is.na(filldata$steps[i])==TRUE) {
      if (filldata$interval[i] %in% mean_intervalsteps$interval) {
        filldata$steps[i] <- mean_intervalsteps$steps[
          which(mean_intervalsteps$interval==filldata[i,3])]
      }
    }
}
# show the head of new data set
head(filldata)
filldatesteps <- melt(filldata[,1:2], id=c("date"),na.rm=TRUE)
mean_filldatesteps<-dcast(filldatesteps, date~variable,mean)
hist(mean_filldatesteps$steps)
# the new mean total number of steps taken per day
mean(mean_filldatesteps$steps)
# the new median total number of steps taken per day
median(mean_filldatesteps$steps) 
```
These values **DO NOT** differ from the estimates from the first part of the assignment. 
Thus, there is **NO BIG IMPACT** of imputing missing data on the estimates of the total daily number of steps.

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
filldata$date <- weekdays(as.Date(filldata$date))
for (i in 1:length(filldata$date)) {
  if (filldata$date[i] != "Sunday" & filldata$date[i] != "Saturday") {
    filldata$date[i] <- "weekday"
  } else {
    filldata$date[i] <- "weekend"
  }
}
# reshape the data to calculate the average steps
library(reshape2)
filldaysteps <- melt(filldata, id=c("date","interval"),na.rm=TRUE)
mean_filldaysteps<-dcast(filldaysteps, interval+date~variable,mean)
head(mean_filldaysteps)
# Make a panel plot 
library(lattice)
mean_filldaysteps <- transform(mean_filldaysteps, date=factor(date))
xyplot(steps~interval|date,data=mean_filldaysteps,
       layout=c(1,2),type="l",ylab="number of steps")

```